\vspace{-15mm}\fontsize{21pt}{10pt}\selectfont\textbf{Wordy Manuscript Title: \\Even Wordier Manuscript Subtitle\thanks{Draft manuscript. Please do not cite without the authors' permission.}}

### 2021-12-13 sections/methods.tex
Change: Update datasets subsection in methods section

  \subsection*{Datasets}

  \vspace{-5mm}
  The study uses three kinds of 16S rRNA sequencing datasets: real datasets, mock datasets and synthetic datasets.
  Real datasets are collections of sequencing reads obtained from naturally occurring microbial community samples.
  The current study used healthy stool samples from a fecal microbiome transplant study~\cite{Kang2017} and healthy saliva samples from a periodontal disease study~\cite{Chen2018} as real datasets for analysis.
  % TODO: How many samples are there for each and average number of reads.
  The mock community 16S datasets are real sequencing data obtained for artificially assembled collections of species in known proportions.
  The mock datasets used for this study, obtained from mockrobiota~\cite{Bokulich2016}, are labelled mock4, mock12 and mock16.
  The mock4 community is composed of 21 bacterial strains.
  Two replicate samples from mock4  contain all species in equal abundances, and two additional replicate samples contain the same species in unequal abundances.
  The mock12 community is composed of 27 bacterial strains that include closely related taxa with some pairs having only one to two nucleotide difference from another.
  The mock16 community is composed of 49 bacteria and 10 archea, all represented in equal amount.
  The synthetic datasets were generated using an artificial read simulator called ART~\cite{Huang2012}.
  Three different microbial composition profiles were used as input; reads were generated using a soil and water microbiome composition profiles from the \ac{emp}~\cite{Thompson2017} and healthy gut microbiome project from the fecal microbiome transplant study~\cite{Kang2017}.
  % TODO: Create a section in the supplementary that describes what the original profiles were and how we used ART (what settings) to generate the synthetic data.
  The reads are simulated using the NCBI RefSeq database as the reference sequence pool and the `art\_illumina' sequence profile with a mutation rate of 2\%.
  The scripts used to generate the synthetic data are in the scripts folder of the repository (\href{https://github.com/segrelab/MiCoNE-pipeline-paper}{https://github.com/segrelab/MiCoNE-pipeline-paper}).

@@

  The first method, hereby referred to as ``seqtime'', used generalized Lotka-Volterra (gLV) equations to model the microbial community dynamics and utilized the Klemm–Eguı́luz algorithm to generate a clique-based interaction network \cite{rottjersHairballsHypothesesBiological2018}. We used the \href{https://github.com/hallucigenia-sparsa/seqtime}{seqtime} R package to simulate communities with number of species (\(N\)) varying from 10 to 150 and number of samples (\(S\)) varying from 50 to 500. The abundance values of the species in the community at the end of the simulation time are used to create the OTU table.

@@

  For each method, we generated the OTU table depicting the abundances of species and used this to generate association matrices using MiCoNE. The interaction matrix was used as the source of expected (true) interactions and the associations predicted using MiCoNE were the source of predicted interactions. We evaluated the precision and sensitivity of the associations predicted by the individual network inference methods as well as the consensus (Figure XXX).

>>>>

  \subsection*{16S rRNA sequencing Datasets}

  \vspace{-5mm}
    The study uses two kinds of 16S rRNA sequencing datasets: biological datasets and mock datasets.
    Biological datasets are collections of sequencing reads obtained from naturally occurring microbial community samples.
    The current analysis used stool samples from a fecal microbiome transplant study of autism~\cite{Kang2017} as biological datasets.
    The study was composed of multiple sequencing runs and those runs that contained paired end reads (runs 2, 3 and 4), were downloaded from Qiita~\cite{qiita} from https://qiita.ucsd.edu/study/description/10532 and used as input sequences for the \ac{micone} pipeline.
    % TODO: How many samples are there for each and average number of reads.
    Sequences from both control (x samples containing xx reads) and autistic (y samples containing yy reads) patients were included in the analyses but the groups were analyzed independently in the pipeline.
    The mock community 16S datasets are sequencing data obtained for artificially assembled collections of species in known proportions.
    The mock datasets used for this study, obtained from mockrobiota~\cite{Bokulich2016}, are labelled mock4, mock12 and mock16.
    The mock4 community is composed of 21 bacterial strains.
    Two replicate samples from mock4 contain all species in equal abundances, and two additional replicate samples contain the same species in unequal abundances.
    The mock12 community is composed of 27 bacterial strains that include closely related taxa with some pairs having only one to two nucleotide difference from another.
    The mock16 community is composed of 49 bacteria and 10 archea, all represented in equal amount.

@@

  The first method, hereby referred to as ``seqtime'', used generalized Lotka-Volterra (gLV) equations to model the microbial community dynamics and utilized the Klemm–Eguı́luz algorithm to generate a clique-based interaction network~\cite{rottjersHairballsHypothesesBiological2018}.
  We used the \href{https://github.com/hallucigenia-sparsa/seqtime}{seqtime} R package to simulate communities with number of species ($N$) varying from 10 to 150 (10, 25, 50, 100, 150 and 200).
  The initial species concentrations are randomly sampled from a Poisson distribution and the simulation is rerun to generate a number of samples ($S$) varying from 50 to 500 (50, 100, 200, 500) for different communities.
  The abundance values of the species in the community at the end of the simulation time are used to create the OTU table.

@@

  For each method, we generated the OTU table depicting the abundances of species and used this to generate association networks using \ac{micone}.
  The interaction matrix was used as the source of expected (true) interactions and the associations predicted using \ac{micone} were the source of predicted interactions.
  Finally, for each dataset we evaluated the precision and sensitivity of the associations predicted by the individual network inference methods as well as the consensus (Figure \ref{fig:figure7}).
